---
title: "Statistical Model To Predict The Weight Of Newborns"
author: "Enrico Michelon"
output:
    pdf_document:
        toc: true
        toc_depth: 3
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
library(grid)
library(gridExtra)
library(car)

getwd()
#setwd("/home/enrico/statistica/statistical_model_to_predict_the_weight_of_newborns/src")

gini.index <- function(x)
{
  fi2 = (table(x)/length(x))^2
  J = length(table(x))
  gini_index = (1-sum(fi2))/((J-1)/J)
  return(gini_index)
}

ggplot_density.plot <- function(dataset, variable, x_label)
{
  return (ggplot(data = dataset, aes(x = variable)) +
            geom_density(color = "blue") +
            geom_vline(xintercept = quantile(variable, 0.25), color = "red") +
            geom_vline(xintercept = quantile(variable, 0.75), color = "red") +
            labs(x = x_label, y = "Density") +
            theme_minimal()+
            theme(panel.grid.major = element_blank(),
                panel.grid.minor = element_blank(),
                axis.line = element_line(color = "black"))
    )
}

compute.statistics <- function(observation)
{
  obj.summary <- round(summary(observation), digits = 2)
  obj.std_dev <- list("std-dev"=round(sd(observation), digits = 2))
  obj.IQR <- list("IQR"=round(IQR(observation), digits = 2))
  obj.skewness <- list("skewness"=round(moments::skewness(observation), digits = 2))
  return(c(obj.summary, obj.std_dev, obj.IQR, obj.skewness))
  
}

figure.Number = 0

```

## Introduction (Point 2)

This project concerns the creation of a statistical model to predict the
weight of newborns. Our objective is to create a statistical model given
the *neonati.csv* dataset that can be extended to the entire population.
 In particular we want to find out whether it is possible to predict
the infant's weight at birth given all the other variables. In
particular, one wants to study a relationship with the mother's
variables, to find out whether or not these have a not have a
significant effect, such as the potentially harmful effect of smoking.
The length and diameter of the infant's skull are also used because they
can be estimated already from ultrasound scans, but in general they
could also serve as control variables.

```{r dataframe_table, results='asis', echo=FALSE}
newborns_dataset <- read.csv("data/neonati.csv")
knitr::kable(head(newborns_dataset), type = "latex", align='c', booktabs = TRUE,
    caption = "Dataset first rows")
```

## Dataset (Points 1, 2, 3)

The dataset is composed by 2500 samples and, studying its first rows, we
can distinguish 10 variables: Anni.madre, N.gravidanze, Fumatrici,
Gestazione, Peso, Lunghezza, Cranio, Tipo.parto, Ospedale and Sesso.

### Anni.madre

Anni.madre is a quantitative variable on radio scale. In the dataset we
have at least two outlayers, which can be found at rows 1152 and 1380,
and report an age of 1 and 0, respectively. Computing position measures
and standard deviation excluding those rows, we obtain:

::: {align="center"}
```{r anni_madre_summary, results='asis', echo=FALSE}
min_val <- min(newborns_dataset$Anni.madre)
line <- which(newborns_dataset$Anni.madre == min_val)
anni_madre_without_outlayers <- newborns_dataset$Anni.madre[
    -c(line, which(newborns_dataset$Anni.madre == 1))]

sigma.anni_madre <- round(sd(anni_madre_without_outlayers), 2)

df_sigma <- data.frame(std.dev = sigma.anni_madre)

knitr::kable(t(c(round(summary(anni_madre_without_outlayers), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Anni.madre")
```
:::

### N.gravidanze

N.gravidanze is a quantitative variable on ratio scale, which represents
the number of pregnancies per mother. In Table 3 position measures and
standard deviation for the variable are shown. We can see that mean
and standard deviation and third interquartile are around 1 (0.98, 1.28
and 1 respectively), while the maximum reaches a value of 12.

::: {align="center"}
```{r n_gravidanze_summary, results='asis', echo=FALSE}

sigma.n_gravidanze <- round(sd(newborns_dataset$N.gravidanze), 2)

df_sigma <- data.frame(std.dev = sigma.n_gravidanze)

knitr::kable(t(c(round(summary(newborns_dataset$N.gravidanze), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Anni.madre")

```
:::

We can look now at the distribution of *N.gravidanze*. From Figure
`r figure.Number <- figure.Number + 1; figure.Number` we can notice that
it is a normal positive skewed distribution, with mean 0.98 and standard
deviation of 1.28.

```{r n_gravidanze_density, results='asis', echo=FALSE, fig.align='center', fig.height=3.5, fig.width=6}

density_plot <- ggplot(data = newborns_dataset, 
        aes(x = N.gravidanze)) +
  geom_bar(fill = 'seagreen2') +
  ylab("Mothers") + 
  xlab("Pregnancy number") +
  scale_x_continuous(breaks = seq(0, 12, 1)) +
  theme_minimal() +
  theme(plot.title = element_text(hjust = 0.5),
    panel.grid.major = element_blank())

caption_text <- sprintf("Figure %d: N.gravidanze distribution", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(density_plot, bottom = caption_text))
```

### Fumatrici

Fumatrici is a qualitative encoded variable on nominal scale, with
values 0 and 1. Value 0 means that the mother is not a smoker, while
mothers with "Fumatrici" value of 1 means she is a smoker. 

```{r fumatrici_summary, echo=FALSE}
fumatrici_gini_coeff <- round(gini.index(newborns_dataset$Fumatrici), 2)
not_smoker_percentage <- table(newborns_dataset$Fumatrici)
```

Computing the Gini coefficient for the variable, we obtain a value of
`r fumatrici_gini_coeff`, which means that the ditribution is not much
equally distributed. In Figure
`r figure.Number <- figure.Number + 1; figure.Number`, the distribution
of variable "Fumatrici" is shown.

```{r fumatrici_graphic, results='asis', echo=FALSE, fig.align='center', fig.height=3, fig.width=5}
smoker_plot <- ggplot(data = newborns_dataset) +
  geom_bar(aes(y = Fumatrici, 
               fill = factor(Fumatrici)), width = 0.45,
           position = "identity") +
  xlab("Mothers") +
  ylab("Smoker (yes or no)") + 
  scale_fill_manual("Smoker (0 = No, 1 = Yes):",
                    values = c("0" = 'lightgreen', "1" = 'indianred3')) +
  scale_y_continuous(breaks = seq(0, 1, 1) ) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.grid.major = element_blank())

caption_text <- sprintf("Figure %d: Smoker vs Not Smoker", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(smoker_plot, bottom = caption_text))
```

### Gestazione

Gestazione is a quantitative variable on ratio scale, measured in weeks,
with position and standard deviation measures that can be seen in Table
5.  As we expect, the mean value is around 40 weeks (38.98), with a
low standard deviation of 1.87 weeks.

::: {align="center"}
```{r gestazione_summary, results='asis', echo=FALSE}

sigma.gestazione <- round(sd(newborns_dataset$Gestazione), 2)

df_sigma <- data.frame(std.dev = sigma.gestazione)   

knitr::kable(t(c(round(summary(newborns_dataset$Gestazione), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Gestazione")
```
:::

### Peso

Peso is a quantitative variable on ratio scale, and represents the
weight of newborns in grams. Position measures and standard deviation
are observable in Table 6.

::: {align="center"}
```{r peso_summary, results='asis', echo=FALSE}

sigma.Peso <- round(sd(newborns_dataset$Peso), 2)

df_sigma <- data.frame(std.dev = sigma.Peso)   

knitr::kable(t(c(round(summary(newborns_dataset$Peso), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Peso")
iqr <- IQR(newborns_dataset$Peso)

```
:::

It is interesting to note that "Peso" has in IQR value of 630 grams,
while the range is of 4100. This can been explained studying the graphic
on Figure `r figure.Number <- figure.Number + 1; figure.Number`. The
distribution is negatively skewed, with very long tails, specially on
the left, making a large different between IQR (represented by red line
on the graphic) and range. The distribution is not a Normal
distribution, as we will see later observing the Q-Q Plot.

```{r peso_density, results='asis', fig.align='center', echo=FALSE, fig.width=5, fig.height=4}
par(mgp = c(2, 1, 0))
peso_density.plot <- plot(density(newborns_dataset$Peso), 
            col = "blue", main = "", 
            xlab = "Newborns weight")
Q1 <- quantile(newborns_dataset$Peso, .25)
Q3 <- quantile(newborns_dataset$Peso, .75)
abline(v = Q1, col = "red")
abline(v = Q3, col = "red")

caption_text <- sprintf("Figure %d: Peso distribution", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(peso_density.plot, bottom = caption_text))

skew <- moments::skewness(newborns_dataset$Peso)

```

### Lunghezza

Lunghezza is a quantitative variable on ratio scale. Represents the
length of the newborn in millimeters. For this variable we expect a
behaviour quite similar to variable "Peso".

::: {align="center"}
```{r Lunghezza_summary, results='asis', echo=FALSE}

sigma.Lunghezza <- round(sd(newborns_dataset$Lunghezza), 2)

df_sigma <- data.frame(std.dev = sigma.Lunghezza)   

knitr::kable(t(c(round(summary(newborns_dataset$Lunghezza), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Lunghezza")
```
:::

```{r Lunghezza_density, results='asis', echo=FALSE, fig.align='center', fig.height=3}

lunghezza_density.plot <- ggplot_density.plot(newborns_dataset, newborns_dataset$Lunghezza, "Lunghezza")

peso_density.plot <- ggplot_density.plot(newborns_dataset, newborns_dataset$Peso, "Peso")
figure.Number <- figure.Number + 1
caption_text <- sprintf("Figure %d: Lunghezza distribution vs. Peso distribution", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(lunghezza_density.plot, peso_density.plot, bottom = caption_text, ncol = 2))


skew <- moments::skewness(newborns_dataset$Peso)
skew <- moments::skewness(newborns_dataset$Lunghezza)
```

### Cranio

Cranio is a quantitative variable on ratio scale. Represents the cranium
diameter of the newborn in millimeters.  In Table 8, we can observe
position measures and standard deviation for variable "Cranio".

::: {align="center"}
```{r cranio_summary, results='asis', echo=FALSE}

sigma.Cranio <- round(sd(newborns_dataset$Cranio), 2)

df_sigma <- data.frame(std.dev = sigma.Cranio)   

knitr::kable(t(c(round(summary(newborns_dataset$Cranio), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Cranio")
```
:::

### Tipo.parto

Tipo.parto is a qualitative variable on nominal scale. Represents the
type of childbirth, and is encoded with "Nat" (natural) and "Ces"
(caesarean). For a qualitative variable we can compute Gini coefficient,
which provides a measure of heterogeneity. Gini coefficient for
"Tipo.parto" is 0.83, indicating a very high heterogeneity.

```{r tipo_parto_gini, include=FALSE}
gini_index.Parto <- gini.index(newborns_dataset$Tipo.parto)
```

### Ospedale

Ospedale is a qualitative variable on nominal scale. It represents three
different hospitals, and is coded in "osp 1", "osp 2" and "osp 3". The
Gini coefficient is very close to 1 which means that the variable is
almost heterogeneous.

```{r ospedale_gini, include=FALSE}
gini_index.Ospedale <- gini.index(newborns_dataset$Ospedale)
```

### Sesso

Sesso is a qualitative variable on nominal scale. Represents the gender
of the newborn, and is coded with "M" for males and "F" for females. The
Gini coefficient is very close to 1 which means that the variable is
almost heterogeneous. In Figure
`r figure.Number <- figure.Number + 1; figure.Number` we can see the
distribution of "Sesso".

```{r sesso_gini, include=FALSE}
gini_index.Sesso <- gini.index(newborns_dataset$Sesso)
```

```{r sesso_graphic, results='asis', echo=FALSE, fig.align='center', fig.height=3.5, fig.width=5}
counts <- newborns_dataset %>%
  count(Sesso)

sesso_plot <- ggplot(data = newborns_dataset) + 
  geom_bar(aes(y = Sesso, fill = factor(Sesso)), width = 0.35, position = "identity") +
  geom_text(data = counts, aes(x = n, y = Sesso, label = n, group = Sesso), 
            position = position_dodge(width = 0.35), angle=-90, vjust = -0.5) +
  scale_fill_manual("Gender:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
  theme_minimal() +
  theme(legend.position = "bottom",
        panel.grid.major = element_blank())

caption_text <- sprintf("Figure %d: Gender distribution", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(sesso_plot, bottom = caption_text))
```

## Descriptive statistics (Points 3, 5)

We want to introduce now a short descriptive analysis of the dataset.
For our porpuses, we divide frequency distribution of Anni.madre into
classes, and we show the distribution in Figure
`r figure.Number <- figure.Number + 1; figure.Number`.

```{r anni_madre_cl_distr, results='asis', fig.align='center', echo=FALSE, fig.height=4.5, fig.width=4}
Anni.madre_cl <- cut(newborns_dataset$Anni.madre, 
                 breaks = c(min(newborns_dataset$Anni.madre), 16, 21, 26, 31, 36, max(newborns_dataset$Anni.madre)+1),
                 include.lowest = TRUE, right = FALSE)

N <- length(Anni.madre_cl)
ni <- table(Anni.madre_cl)
fi <- ni/N
Ni <- cumsum(ni)
Fi <- Ni/N

distr_freq.Anni.madre_cl <- as.data.frame(cbind(ni, fi, Ni, Fi))

anni_madre_cl_distr <- ggplot(distr_freq.Anni.madre_cl, 
                    aes(x = rownames(distr_freq.Anni.madre_cl), y = ni, group = 1)) +
                geom_bar(stat = "identity", fill = "tan2", width = 0.75) +
                geom_line(col = "coral3") +
                geom_point(col = "coral3", size = 2) +
                geom_text(data = distr_freq.Anni.madre_cl, 
                    aes(x = rownames(distr_freq.Anni.madre_cl), y = ni, label = ni, group = 1, vjust = -0.5)) +
                scale_y_continuous(breaks = seq(0, 950, 100)) +
                labs(x = "Anni.madre", y = "Absolute frequency") +
                theme_minimal() +
                theme(panel.grid.minor = element_blank())

caption_text <- sprintf("Figure %d: Frequency distribution of \"Anni.madre\" class", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(anni_madre_cl_distr, bottom = caption_text))
```

Once obtained the distribution in classes of Anni.madre, we are
interested in observing the trend of number of pregnancies per class of
ages.

```{r descr_stats_01, results='asis', fig.align='center', echo=FALSE, fig.height=3, fig.width=4.5}
dataset_with_class_anni_madre <- cbind(newborns_dataset, Anni.madre_cl)

anni_madre_cl_distr <- ggplot(data = dataset_with_class_anni_madre,
                        aes(x = Anni.madre_cl, fill = factor(N.gravidanze))) +
                        geom_bar(position = "fill") +
                        labs(x = "Mothers age", y = "N. of pregnancies", fill = "N. of Pregnancies:") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

figure.Number <- figure.Number + 1
caption_text <- sprintf("Figure %d: N. of pregnancies per class of ages", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(anni_madre_cl_distr, bottom = caption_text))

```

As we can see in Figure `r figure.Number` as we could expect, the number
of pregnancies has an increasing trend. Moreover, in the classes of ages
between 31 and 35, there is still more than 25% of mothers, which are
waiting the first baby. Another interesting result is obtained observing
the class of age 0-15, where a consistent percentage of mothers are
waiting for the second baby. However, the latter statistic is somewhat
misleading, as the sample of mothers between the ages of 0 and 15 is
obviously very small.  Computation of this statistic in general does
not give us particular information. Therefore, we can proceed by
observing the type of childbirth, subordinate to the age and the fact
that mother smokes or not. 

From Figure `r figure.Number <- figure.Number +1; figure.Number`, we can note that
until the range 26-30, there seems to be no impact for smoker mothers
with respect to non smoker mothers. On the other hand, as age passing,
we note that the number of caesarean births decreases if the mother is a
smoker, which is really strange. However we wanted to show you this
statistic, to remember that great care must be taken when analysing
data. In fact, this statistic is highly altered by the fact that the
number of smoking mothers, as shown in Figure 2, is only a small
percentage compared to the number of non-smoking mothers.

```{r descr_stats_02, results='asis', fig.align='center', echo=FALSE, fig.height=6.5, fig.width=5}

type_of_childbird <- dataset_with_class_anni_madre %>%
    group_by(Anni.madre_cl, Fumatrici) %>%
    reframe(Tipo.parto = Tipo.parto)

anni_madre_cl_distr <- ggplot(data = type_of_childbird,
                              aes(x = Fumatrici, fill = Tipo.parto, group = factor(Tipo.parto))) +
                        geom_bar(position = "fill", width = 0.5) +
                        scale_fill_manual("Byrth type:", 
                              values = c("Nat" = "darkcyan",  
                                         "Ces" = "coral"))+
                        scale_x_continuous(breaks = seq(0,1,1)) +
                        labs(x = "Smoker (no/yes)", y = "Proportion of natural and caesarean births") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
                        facet_wrap(~ Anni.madre_cl, ncol = 1, strip.position = "right")

caption_text <- sprintf("Figure %d: Birth type per class of ages, subordinates to smoker", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(anni_madre_cl_distr, bottom = caption_text))
```

We want now to analyse the skull circumference with respect to pregnancy
and class of ages of mothers. From Figure
`r figure.Number <- figure.Number +1; figure.Number`, we can observe
that the class of ages does not significantly affect the skull
circumference, whereas, as expected, weeks of gestation significantly
affect skull size. Later on, we will analyse how and whether gender
affects the size of the fetus' skull.

```{r descr_stats_03, results='asis', fig.align='center', echo=FALSE, fig.height=7.5, fig.width=5.5}
gestazione <- dataset_with_class_anni_madre %>%
    group_by(Anni.madre_cl, Gestazione) %>%
    reframe(Cranio = Cranio)

anni_madre_cl_distr <- ggplot(data = gestazione,
                              aes(x = Gestazione, fill = Cranio, group = factor(Cranio))) +
                        geom_bar(position = "fill", width = 0.5) +
                        labs(x = "Weeks of pregnancy", y = "Skull circumference proportion", fill = "Skull circumference:") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom") +
                        facet_wrap(~ Anni.madre_cl, ncol = 1, strip.position = "right")

caption_text <- sprintf("Figure %d: Skull circumference, subordinates to weeks of pregnancy and class of ages", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(anni_madre_cl_distr, bottom = caption_text))
```

In Figure `r figure.Number <- figure.Number +1; figure.Number`, we can
observe the different behaviour of weight based on the gender of the
newborns. Since the statistic contained a lot of samples, we have only
shown gestations in the interquartile range in the graph. Later on, we
will analyse the full dataset, through a boxplot.

```{r descr_stats_04, results='asis', fig.align='center', echo=FALSE, fig.height=3.75, fig.width=4.5}
Q1 <- quantile(newborns_dataset$Gestazione, .25)
Q3 <- quantile(newborns_dataset$Gestazione, .75)

options(dplyr.summarise.inform = FALSE)
gestazione <- newborns_dataset %>%
    filter(Gestazione >= quantile(Gestazione, 0.25) & Gestazione <= quantile(Gestazione, 0.75))  %>%
    group_by(Sesso, Gestazione) %>%
    summarise(Peso = round(mean(Peso)))
    
options(dplyr.summarise.inform = TRUE)
peso_conditioned_sesso <- ggplot(data = gestazione,
                              aes(x = Gestazione, y = Peso, colour = Sesso, group = factor(Sesso))) +
                        geom_line(lwd = 2) +
                        geom_point(size=3) +
                        scale_color_manual("Gender:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
                        scale_x_continuous(breaks = seq(38,40,1)) +
                        labs(x = "Weeks of pregnancy", y = "Weight") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

caption_text <- sprintf("Figure %d: Weight of newborns per gender and weeks of pregnancy", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(peso_conditioned_sesso, bottom = caption_text))

```

In Figure `r figure.Number <- figure.Number +1; figure.Number` instead,
we can observe the different behaviour of length of newborns,
considering a caesarean birth or natural birth, depending on the weeks
of pregnancy. This statistic has been computed with the idea to check
whether the caeserean birth, has been induced by the fetus dimensions,
for example in case of fetal macrosomia. Clearly, this statistic no
longer gives us the reason for the choice of type of birth, however, it
is interesting to note that in most cases, the length of the fetus born
by caesarean birth was greater than or equal to the length of the fetus
born by natural childbirth.

```{r descr_stats_05, results='asis', fig.align='center', echo=FALSE, fig.height=3.75, fig.width=4.5}
options(dplyr.summarise.inform = FALSE)
gestazione <- dataset_with_class_anni_madre %>%
    group_by(Tipo.parto, Gestazione) %>%
    reframe(Lunghezza = mean(Lunghezza))
    
options(dplyr.summarise.inform = TRUE)

anni_madre_cl_distr <- ggplot(data = gestazione,
                              aes(x = Gestazione, y = Lunghezza, colour = Tipo.parto, group = factor(Tipo.parto))) +
                        geom_line() +
                        scale_color_manual("Byrth type:", 
                              values = c("Nat" = "darkcyan",  
                                         "Ces" = "coral"))+
                        geom_point(size=2) +
                        labs(x = "Weeks of pregnancy", y = "Child length") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

caption_text <- sprintf("Figure %d: Length of newborns, based on birth-type and gestation", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(anni_madre_cl_distr, bottom = caption_text))
```

Finally in Figure `r figure.Number <- figure.Number +1; figure.Number`,
we can observe the differences of weight, length, skull circumference
and type of childbirth, between genders. It is interesting to note the
total independence between the sex of the fetus and the type of birth.
On the other hand, there is a difference between gender observations
concerning physical characteristics. The most significant difference can
be noticed in weight of newborns.

```{r measures_comparison_between_genders, fig.align='center', echo=FALSE}
dataframe_males <- newborns_dataset %>%
  filter(Sesso == 'M') %>%
  reframe(Sesso, Peso, Lunghezza, Cranio)

dataframe_females <- newborns_dataset %>%
  filter(Sesso == 'F') %>%
  reframe(Sesso, Peso, Lunghezza, Cranio)

M_stats.peso <- compute.statistics(dataframe_males$Peso)
M_stats.lunghezza <- compute.statistics(dataframe_males$Lunghezza)
M_stats.cranio <- compute.statistics(dataframe_males$Cranio)
F_stats.peso <- compute.statistics(dataframe_females$Peso)
F_stats.lunghezza <- compute.statistics(dataframe_females$Lunghezza)
F_stats.cranio <- compute.statistics(dataframe_females$Cranio)
```

With the support of Figure `r figure.Number` and Tables 8 and 9, we can
see that IQR ranges for variable "Peso", are equal, while the total
range is much wider for females, meaning that outlayers are move further
away from the average value. Moreover this means that also the standard
deviation is higher for females with respect to males.

```{r print_measures_comparison_between_genders_peso, fig.align='center', echo=FALSE}
knitr::kable(t(M_stats.peso),
    type="html", align='c', 
    caption = "Statistics for Peso (males)")

knitr::kable(t(F_stats.peso),
    type="html", align='c', 
    caption = "Statistics for Peso (females)")
```

Differently in Tables 10 and 11, we can see that IQR ranges, standard
deviation, mean and median for variable "Lunghezza", are quite similar.
However, the total range is still a bit wider in females.

```{r print_measures_comparison_between_genders_lunghezza, fig.align='center', echo=FALSE}
knitr::kable(t(M_stats.lunghezza),
    type="html", align='c', 
    caption = "Statistics for Lunghezza (males)")
knitr::kable(t(F_stats.lunghezza),
    type="html", align='c', 
    caption = "Statistics for Lunghezza (females)")
```

Finally for variable "Cranio" (Tables 12 and 13, Figure
`r figure.Number`c), shows results comparable to the variable
"Lunghezza".

```{r print_measures_comparison_between_genders_cranio, fig.align='center', echo=FALSE}
knitr::kable(t(M_stats.cranio),
    type="html", align='c', 
    caption = "Statistics for Cranio (males)")
knitr::kable(t(F_stats.cranio),
    type="html", align='c', 
    caption = "Statistics for Cranio (females)")


```

```{r descr_stats_06, results='asis', fig.align='center', echo=FALSE, fig.height=7, fig.width=6}

options(dplyr.summarise.inform = FALSE)
dataframe_gender <- newborns_dataset %>%
    group_by(Sesso) %>%
    reframe(Peso, Lunghezza, Cranio, Tipo.parto)
    
options(dplyr.summarise.inform = TRUE)
peso_conditioned_sesso <- ggplot(data = dataframe_gender) +
                        geom_boxplot(aes(x = Sesso, y = Peso, fill = Sesso)) +
                        scale_fill_manual("Gender:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
                        labs(x = "Gender", y = "Weight") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

lunghezza_conditioned_sesso <- ggplot(data = dataframe_gender) +
                        geom_boxplot(aes(x = Sesso, y = Lunghezza, fill = Sesso)) +
                        scale_fill_manual("Sex:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
                        labs(x = "Gender", y = "Length") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

cranio_conditioned_sesso <- ggplot(data = dataframe_gender) +
                        geom_boxplot(aes(x = Sesso, y = Cranio, fill = Sesso)) +
                        scale_fill_manual("Sex:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
                        labs(x = "Gender", y = "Skull circumference") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")


tipo_parto_conditioned_sesso <- ggplot(data = dataframe_gender,
                              aes(x = Tipo.parto, fill = Sesso)) +
                        geom_bar(position = "fill") +
                        scale_fill_manual("Sex:", values = c("M" = 'lightblue', "F" = 'lightpink')) +
                        labs(x = "Childbird", y = "Gender") +
                        theme_minimal() +
                        theme(panel.grid.minor = element_blank(), legend.position = "bottom")

caption_text <- sprintf("Figure %d: Gender comparisons for weight, length, skull circumference and type of childbirth", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(peso_conditioned_sesso, lunghezza_conditioned_sesso, 
        cranio_conditioned_sesso, tipo_parto_conditioned_sesso, bottom = caption_text, 
        ncol = 2, nrow = 2))

```

## Hypothesis testing (Points 4 and 6)

We now want to understand whether the sample of infants extracted
correctly represents the population. In particular we want to test the
hypothesis that the mean of weight and length of newborns is the same as
the population. From *ospedalebambinogesu.it* we got that mean of weight
and length of population of fetus is 3300 grams and 50 cm respectively.
As we said on paragraph Peso, distribution of "Peso" and "Lunghezza" are
not normal. In Figure
`r figure.Number <- figure.Number +1; figure.Number` we can see the Q-Q
Plots for the two variables, which show a strong deviation from
normality.

```{r qqplot_peso_lunghezza, results='asis', fig.align='center', echo=FALSE, fig.height=3, fig.width=6}

qq_plot_peso <- ggplot(newborns_dataset, aes(sample = Peso)) +
  stat_qq(col="coral1", shape = 1, size = 2) +
  stat_qq_line(col="coral1") +
  ggtitle("Q-Q Plot for Peso") +
  labs(x = "Theoretical Quantiles", y= "Sample Quantiles") +
  theme_minimal() +
  theme(axis.title=element_text(size=8), axis.text=element_text(size=8), plot.title = element_text(hjust = 0.5, size = 9))

qq_plot_lunghezza <- ggplot(newborns_dataset, aes(sample = Lunghezza)) +
  stat_qq(col="indianred", shape = 1, size = 2) +
  stat_qq_line(col="indianred") +
  ggtitle("Q-Q Plot for Lunghezza") +
  labs(x = "Theoretical Quantiles", y= "Sample Quantiles") +
  theme_minimal() +
  theme(axis.title=element_text(size=8), axis.text=element_text(size=8), plot.title = element_text(hjust = 0.5, size = 9))

caption_text <- sprintf("Figure %d: Q-Q Plot for Peso and Lunghezza", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(qq_plot_peso, qq_plot_lunghezza, bottom = caption_text, ncol = 2))

```

We decided to compute Student T-test, exploiting the central limit
theorem, and then the one sample Wilcoxon rank sum, to test the
hypothesis reported before. With both test we have the same result,
which is to not reject the null hypothesis for variable Peso, while to
reject the null hypothesis for variable Lunghezza.

```{r test_hyp, results='asis', fig.align='center', echo=FALSE, fig.height=3, fig.width=7}

#TEST T con criterio SOGLIA
peso_t_test_result <- t.test(newborns_dataset$Peso, 
       mu = 3300,
       conf.level = 0.95, 
       alternative = "two")
#non rifiuto l'ipotesi nulla, perché la mia media è all'interno dell'intervallo di confidenza.

peso_wilcox_test_result <- wilcox.test(newborns_dataset$Peso, mu = 3300, conf.level = 0.95, alternative = "two.sided")


Lunghezza_distr_IQR <- newborns_dataset %>%
  filter(Lunghezza >= quantile(Lunghezza, 0.25) & Lunghezza <= quantile(Lunghezza, 0.75))

lunghezza_t_test_result <- t.test(newborns_dataset$Lunghezza, 
       mu = 500,
       conf.level = 0.95, 
       alternative = "two")
lunghezza_wilcox_result <- wilcox.test(newborns_dataset$Lunghezza, mu = 500, conf.level = 0.95, alternative = "two.sided")
lunghezza_wilcox_result_IQR <- wilcox.test(Lunghezza_distr_IQR$Lunghezza, mu = 500, conf.level = 0.95, alternative = "two.sided")

par(mgp = c(2, 1, 0), mfrow=c(1, 2))

t_test_plot_1 <- plot(density(rt(10000000, 2499)), main = "", 
            xlab = "Test-t for Peso")
abline(v = qt(0.025, 2499), col = "red")
abline(v = qt(0.975, 2499), col = "red")
abline(v = peso_t_test_result$statistic, col = "black")

t_test_plot_2 <- plot(density(rt(10000000, 2499)), main = "", 
            xlab = "Test-t for Lunghezza", xlim=c(-11,11))
abline(v = qt(0.025, 2499), col = "red")
abline(v = qt(0.975, 2499), col = "red")
abline(v = lunghezza_t_test_result$statistic, col = "black")
figure.Number <- figure.Number +1
caption_text <- sprintf("Figure %d: Decision rule for Peso and Lunghezza", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(t_test_plot_1, t_test_plot_2, bottom = caption_text, ncol=2))

```

On Figure `r figure.Number`, we can see the decision rule for Peso and
Lunghezza. In case of Peso we are inside the acceptation zone while in
case of Lunghezza we are inside the rejecting zone. Moreover for
variable Peso we have a P-value that is `r peso_t_test_result$p.value`
which is higher than 0.05, using a Student T-test, while it is
`r peso_wilcox_test_result$p.value` using Wilcoxon. For variable
Lunghezza instead, the corresponding P-value are
`r lunghezza_t_test_result$p.value` and
`r lunghezza_wilcox_result$p.value`. Even if mean value of lenght of our
dataset differs from the global mean value by just 4 millimeters, the
null hypothesis has been rejected. This could be caused by the low
standard deviation (and consequently variance), which cause the sample
to be distorted but precise, and the proof is given by another Wilcoxon
test for Lunghezza, considering only measures of Lunghezza inside the
interquartile range, which provide the same result. 

``` {r test_hyp_hospital, results='asis', fig.align='center', echo=FALSE, include=FALSE}
chi_square_test_result <- chisq.test(newborns_dataset$Tipo.parto, newborns_dataset$Ospedale)

df_osp_1 <- newborns_dataset %>%
  filter(Ospedale == "osp1") %>%
  count(Tipo.parto) %>%
  reframe(Tipo.parto, freq = round(n/sum(n), 2), Ospedale = "osp1")

df_osp_2 <- newborns_dataset %>%
  filter(Ospedale == "osp2") %>%
  count(Tipo.parto) %>%
  reframe(Tipo.parto, freq = round(n/sum(n), 2), Ospedale = "osp2")

df_osp_3 <- newborns_dataset %>%
  filter(Ospedale == "osp3") %>%
  count(Tipo.parto) %>%
  reframe(Tipo.parto, freq = round(n/sum(n), 2), Ospedale = "osp3")

union_df <- rbind(df_osp_1, df_osp_2, df_osp_3)

#test if in some hospitals there are more cesarean than in the others
wilcox_test_result <- pairwise.wilcox.test(union_df$freq, union_df$Ospedale, paired = FALSE, p.adjust.method = "bonferroni")
```

Another point we want to test is **if the percentage of caesarean section is the 
same for all the hospitals, or there is some hospitals in which this percentage 
is higher**.
To do this, we use a chi-square test, which is a test for indipendence between 
two categorical variables. The null hypothesis is that the two variables are 
indipendent, while the alternative hypothesis is that they are not. The result 
of this test is that we can not reject the null hypotesis, obtaining a p-value 
of `r round(chi_square_test_result$p.value, 2)`. The other test we made was a pairwise 
Wilcoxon test, which is a non-parametric test for multiple comparisons. 
The null hypothesis is that the three hospitals have the same percentage of
caesarean section, while the alternative hypothesis is that they are not. The 
result of this test is that we can not reject the null hypotesis, obtaining a 
p-value of 1 comparing hospitals. This means that the percentage of caesarean
section is the same for all the hospitals, and there is no hospital in which
this percentage is higher than the others.

## Part 2: Generalized Linear Models (Point 1)
In this part we will use the dataset `newborns_dataset` to want to build a model
to predict the weight of a newborn, using the other variables of the dataset.
The first thing we want to do is to investigate the relationship between the
weight of a newborn and the other variables. We can investigate this relashionship
both numerically, by using covariance and coefficient of linear correlation and 
graphically by means of scatterplots.
```{r scatterplot, results='asis', fig.align='center', echo=FALSE, fig.height=2.75, fig.width=5}
par(mgp=c(2,1,0))
scat_1 <- scatterplot(Peso ~ Gestazione, 
            data = newborns_dataset, 
            pch = 20, 
            main = "Scatterplot of weight and pregnancy")

scat_2 <- scatterplot(Peso ~ Lunghezza, 
            data = newborns_dataset, 
            smooth = FALSE, 
            pch = 20,
            main = "Scatterplot of weight and length")

scat_3 <- scatterplot(Peso ~ Cranio, 
            data = newborns_dataset, 
            pch = 20,
            smooth = FALSE,
            main = "Scatterplot of weight and skull")

scat_4 <- scatterplot(Peso ~ N.gravidanze, 
            data = newborns_dataset, 
            pch = 20, 
            smooth = FALSE,
            boxplots = FALSE,
            main = "Scatterplot of weight and number of pregnancies")
figure.Number <- figure.Number + 1
caption_text <- sprintf("Figure %d: Scatterplots for weight, and other variables", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(scat_1, scat_2, scat_3, scat_4, bottom = caption_text, ncol = 2))

w_g_cov <- cov(newborns_dataset$Peso, newborns_dataset$Gestazione)
ro_w_g <-cor(newborns_dataset$Peso, newborns_dataset$Gestazione)

w_l_cov <- cov(newborns_dataset$Peso, newborns_dataset$Lunghezza)
ro_w_l <- cor(newborns_dataset$Peso, newborns_dataset$Lunghezza)

w_s_cov <- cov(newborns_dataset$Peso, newborns_dataset$Cranio)
ro_w_s <- cor(newborns_dataset$Peso, newborns_dataset$Cranio)

w_ng_cov <- cov(newborns_dataset$Peso, newborns_dataset$N.gravidanze)
ro_w_ng <- cor(newborns_dataset$Peso, newborns_dataset$N.gravidanze)

```
In Figure ```r figure.Number``` we can see the scatterplots of the weight of a newborn and the other variables.
The scatterplot of the weight and pregnancy length shows a linear relationship, 
with coefficient of linear correlation of `r round(ro_w_g, 2)`.
The scatterplot of the weight and length of the newborn shows a linear relationship,
with a coefficient of linear correlation of `r round(ro_w_l, 2)`.
Similar result we can see in the scatterplot of the weight and the skull of the newborn,
where the linear correlation coefficient is `r round(ro_w_s, 2)`.
On the other hand, the scatterplot of the weight and the number of pregnancies shows
an absence of relationship, with linear correlation coefficient of `r round(ro_w_ng, 2)`.  

One final scatterplot matrix to resume the correlation between "Peso" and other
variables of the dataset can be observed in Figure ```r figure.Number + 1```.
```{r scatterplot_matrix ,  results='asis', fig.align='center', echo=FALSE, fig.height=6, fig.width=6}
scat_matrix <- scatterplotMatrix(~Peso + Gestazione + Lunghezza + Cranio, data = newborns_dataset,pch = 20, main = "Scatterplot matrix")
caption_text <- sprintf("Figure %d: Scatterplot matrix for weight, and variables", figure.Number)
caption_text <- textGrob(caption_text, gp = gpar(fontsize = 9), just = "center")
grid.draw(arrangeGrob(scat_matrix, bottom = caption_text, ncol = 2))
```

### Multiple linear regression model (Point 2)
We want to build a multiple linear regression model to predict the weight of a newborn.
The model we want to build is the following:
$$
Peso = \beta_0 + \beta_1 \cdot Gestazione + \beta_2 \cdot Lunghezza + \beta_3 \cdot Cranio + \beta_4 \cdot N.gravidanze + \beta_5 \cdot Fumatrici + \beta_6 \cdot Sesso + \beta_7 \cdot Ospedale + \epsilon
$$
where $\epsilon \sim N(0, \sigma^2)$. We can build this model by using the function
`lm()` in R. The result of this model is the following:
```{r linear_regression_model, results='asis', fig.align='center', echo=FALSE}
linear_regression_model <- lm(Peso ~ Gestazione + Lunghezza + Cranio + N.gravidanze + Fumatrici + Sesso + Ospedale, data = newborns_dataset)
summary_lm <- summary(linear_regression_model)


```

```{r 3D_scatterplot ,  results='asis', fig.align='center', echo=FALSE, fig.height=6, fig.width=6}
#newborns_dataset$Peso <- as.numeric(newborns_dataset$Peso)
#newborns_dataset$Gestazione <- as.numeric(newborns_dataset$Gestazione)
#newborns_dataset$Lunghezza <- as.numeric(newborns_dataset$Lunghezza)

# Plotting with scatter3d()
#scatter3d(Peso ~ Gestazione + Lunghezza, data = newborns_dataset, 
#               main = "3D scatterplot of weight, pregnancy and length")


```

```{r investigation, results='asis', fig.align='center', echo=FALSE, fig.height=3, fig.width=7}
#parte
```

```{r parte_2, results='asis', fig.align='center', echo=FALSE, fig.height=3, fig.width=7}


```
