---
title: "Statistical Model To Predict The Weight Of Newborns"
author: "Enrico Michelon"
output:
    pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr)
library(ggplot2)
```
## Introduction
This project concerns the creation of a statistical model to predict the weight of newborns. Our objective is to create a
statistical model given the *neonati.csv* dataset that can be extended to the entire population.<br>
```{r dataframe_table, results='asis', echo=FALSE}
newborns_dataset <- read.csv("data/neonati.csv")
knitr::kable(head(newborns_dataset), type = "latex", align='c', booktabs = TRUE, , 
    caption = "Dataset first rows")
```

## Dataset
Studing the first rows of the dataset, we can distinguish 10 variables: Anni.madre, N.gravidanze, Fumatrici, Gestazione, 
Peso, Lunghezza, Cranio, Tipo.parto, Ospedale and Sesso.

### Anni.madre
Anni.madre is a quantitative variable on radio scale. In the dataset we have at least two outlayers, 
which can be found at rows 1152 and 1380, and report an age of 1 and 0, respectively. 
Computing position measures and standard deviation excluding those rows, we obtain:
<div align="center">

```{r anni_madre_summary, results='asis', echo=FALSE}
min_val <- min(newborns_dataset$Anni.madre)
line <- which(newborns_dataset$Anni.madre == min_val)
anni_madre_without_outlayers <- newborns_dataset$Anni.madre[
    -c(line, which(newborns_dataset$Anni.madre == 1))]

sigma.anni_madre <- round(sd(anni_madre_without_outlayers), 2)

df_sigma <- data.frame(std.dev = sigma.anni_madre)

knitr::kable(t(c(round(summary(anni_madre_without_outlayers), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Anni.madre")
```

</div>

### N.gravidanze
N.gravidanze is a quantitative variable on ratio scale. In Table 3 position measures
and standard deviation for the variable are shown.<br>
We can see that mean and standard deviation and third interquartile are around 1 
(0.98, 1.28 and 1 respectively), while the maximum reaches a value of 12.
<div align="center">

```{r n_gravidanze_summary, results='asis', echo=FALSE}

sigma.n_gravidanze <- round(sd(newborns_dataset$N.gravidanze), 2)

df_sigma <- data.frame(std.dev = sigma.n_gravidanze)

knitr::kable(t(c(round(summary(newborns_dataset$N.gravidanze), 2), df_sigma)),
    type="html", align='c', 
    caption = "Position measures and standard deviation for Anni.madre")
```

</div>

We can look now at the distribution of *N.gravidanze*.

<div align="center">

```{r n_gravidanze_density, results='asis', echo=FALSE}


ggplot(data = newborns_dataset, 
        aes(x = N.gravidanze)) +
  geom_density(color = 'blue', fill = 'lightblue') +
  labs(title = 'Continuous Density Plot') +
  scale_x_continuous(breaks = seq(0, 13, 1)) +
  theme_minimal()

```

</div>


### Fumatrici
### Gestazione
### Peso
### Lunghezza
### Cranio
### Tipo.parto
### Ospedale
### Sesso


