---
title: "Statistical Model To Predict The Weight Of Newborns"
author: "Enrico Michelon"
output:
    pdf_document: default
---

```{r setup, include=FALSE}
library(dplyr) 
```
## Introduction
This project concerns the creation of a statistical model to predict the weight of newborns. Our objective is to create a
statistical model given the *neonati.csv* dataset that can be extended to the entire population.<br>
```{r dataframe_table, results='asis', echo=FALSE}
newborns_dataset <- read.csv("data/neonati.csv")
knitr::kable(head(newborns_dataset), type = "latex", align='c', booktabs = TRUE)
```

## Dataset
Studing the first rows of the dataset, we can distinguish 10 variables: Anni.madre, N.gravidanze, Fumatrici, Gestazione, 
Peso, Lunghezza, Cranio, Tipo.parto, Ospedale and Sesso.

### Anni.madre
Anni.madre is a quantitative variable on radio scale. In the dataset we have at least two outlayers, 
which can be found at rows 1152 and 1380, and report an age of 1 and 0, respectively. 
```{r anni_madre_summary, results='asis', echo=FALSE}
summary(newborns_dataset$Anni.madre)
min_val <- min(newborns_dataset$Anni.madre)
#summary return 0 as min value, meaning that there's at least one missing value. 
#So I excluded all lines with 0 "age" to compute position, variance measures and shape measures.
line <- which(newborns_dataset$Anni.madre == min_val)
anni_madre_without_outlayers <- newborns_dataset$Anni.madre[-c(line, which(newborns_dataset$Anni.madre == 1))]

#cat("\\begin{table}[H]\n\\centering\n")
knitr::kable(t(summary(anni_madre_without_outlayers)), format = "latex", align='c')
#cat("\\end{table}\n")

?knitr::kable
```
### N.gravidanze
### Fumatrici
### Gestazione
### Peso
### Lunghezza
### Cranio
### Tipo.parto
### Ospedale
### Sesso


